---
title: "Getting Started with ACSloadR"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with ACSloadR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(eval = FALSE, collapse = TRUE, comment = "#>")
```

ACSloadR provides topic-specific wrappers around `tidycensus::get_acs()`.
Each wrapper adds ACS metadata, parses table semantics, and returns a named
bundle of long data frames.

## Age, sex, race, and ethnicity

```{r}
library(ACSloadR)

age <- get_acs_age(
  year = 2024,
  survey = "acs5",
  geography = "state",
  state = "MA"
)

age$total_population
age$race_ethnicity
```

The components are separate because B01001 and B01001A-I publish different
age bands. ACSloadR does not row-bind or silently harmonize them.

## Employment and occupation

```{r}
employment <- get_acs_employment(2024, "acs5", "county", state = "MA")

employment$employment |>
  dplyr::filter(
    characteristic_type == "age",
    measure == "unemployment_rate"
  )

occupation <- get_acs_occupation(2024, "acs5", "state", state = "MA")

occupation$occupation |>
  dplyr::filter(
    sex == "Female",
    measure == "share_of_sex_employment"
  )

occupation$race_ethnicity |>
  dplyr::filter(
    race_ethnicity == "Black or African American alone",
    sex == "Total",
    measure == "share_of_race_sex_employment"
  )
```

`S2301` rates are published ACS measures. Occupation distribution shares are
derived only where the table provides an unambiguous denominator. Occupation
race and ethnicity tables remain a separate component because ACS publishes
them at less detail than `S2401`, with different source families in the 1-year
and 5-year surveys. Total-sex occupation rows in this component are derived by
summing the published male and female estimates. They retain both variable IDs
in `source_variables` and use the ACS approximate MOE formula for a sum.

## Earnings and commuting

```{r}
earnings <- get_acs_earnings(2024, "acs5", "state", state = "MA")
earnings$earnings

commuting <- get_acs_commuting(2024, "acs5", "state", state = "MA")
commuting$commuting |>
  dplyr::filter(measure == "share_of_table_universe")
```

Every component retains `variable`, `label`, `concept`, `table`, `year`, and
`survey`. Derived shares additionally identify their denominator variable.
