---
title: "Getting Started with ACSloadR"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with ACSloadR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(eval = FALSE, collapse = TRUE, comment = "#>")
```

ACSloadR provides topic-specific wrappers around `tidycensus::get_acs()`.
Each wrapper adds ACS metadata, parses table semantics, and returns a named
bundle of long data frames.

## Age, sex, race, and ethnicity

```{r}
library(ACSloadR)

age <- get_acs_age(
  year = 2024,
  survey = "acs5",
  geography = "state",
  state = "MA"
)

age$total_population
age$race_ethnicity
```

The components are separate because B01001 and B01001A-I publish different
age bands. ACSloadR does not row-bind or silently harmonize them.

## Employment and labor market topics

### Baseline and subject summaries

```{r}
employment <- get_acs_employment(2024, "acs5", "county", state = "MA")

employment$employment |>
  dplyr::filter(
    characteristic_type == "age",
    measure == "unemployment_rate"
  )
```

### Detailed employment status and demographics

```{r}
emp_detail <- get_acs_employment_detail(2024, "acs5", "county", state = "MA")
emp_detail$status
emp_detail$age_sex
emp_detail$race_ethnicity
emp_detail$education
emp_detail$poverty_disability
```

### Work experience, hours, and full-time status

```{r}
work <- get_acs_work_experience(2024, "acs5", "state", state = "MA")
work$hours_weeks
work$full_time_by_age
work$hours_summary
```

### Family and child employment

```{r}
family <- get_acs_family_employment(2024, "acs5", "state", state = "MA")
family$children_parent_status
family$females_with_children
family$family_type_status
family$family_workers
```

### Industry and Class of Worker

```{r}
industry <- get_acs_industry(2024, "acs5", "state", state = "MA")
industry$industry
industry$earnings
industry$industry_by_occupation

cow <- get_acs_class_of_worker(2024, "acs5", "state", state = "MA")
cow$class_of_worker
cow$earnings
cow$class_by_occupation
```

### Occupation and detailed categories

```{r}
occupation <- get_acs_occupation(2024, "acs5", "state", state = "MA")
occupation$occupation
occupation$race_ethnicity

occ_detail <- get_acs_occupation_detailed(2024, "acs5", "state", state = "MA")
occ_detail$occupation
occ_detail$earnings
```

```{r}
earnings <- get_acs_earnings(2024, "acs5", "state", state = "MA")
earnings$earnings

commuting <- get_acs_commuting(2024, "acs5", "state", state = "MA")
commuting$commuting |>
  dplyr::filter(measure == "share_of_table_universe")
```

## Household, Family, & Individual Income & Inequality

```{r}
# Household income brackets and medians
hh_inc <- get_acs_household_income(2024, "acs5", "state", state = "MA")
hh_inc$income_brackets
hh_inc$median_income

# Family income and children presence
fam_inc <- get_acs_family_income(2024, "acs5", "state", state = "MA")
fam_inc$family_brackets
fam_inc$family_medians

# Income composition (earnings, wages, SSI, retirement, SNAP)
inc_types <- get_acs_income_types(2024, "acs5", "state", state = "MA")
inc_types$household_counts

# Gini Index and Quintile shares
ineq <- get_acs_income_inequality(2024, "acs5", "state", state = "MA")
ineq$gini_index
ineq$quintile_shares

# Individual income & earnings by sex and work experience
ind_inc <- get_acs_individual_income(2024, "acs5", "state", state = "MA")
ind_inc$earnings_brackets
ind_inc$earnings_work_exp
```

## School Enrollment, Educational Attainment, & Field of Degree

```{r}
# School enrollment level and detailed grades
enroll <- get_acs_school_enrollment(2024, "acs5", "state", state = "MA")
enroll$enrollment_level
enroll$enrollment_detailed

# Highest educational attainment by age and sex
attain <- get_acs_educational_attainment(2024, "acs5", "state", state = "MA")
attain$attainment_sex
attain$attainment_detailed

# Undergraduate Bachelor's degree field of degree majors
field <- get_acs_field_of_degree(2024, "acs5", "state", state = "MA")
field$field_detailed
field$earnings_by_sex
```

## Commute, Travel Time, Place of Work, & Mode Cross-tabulations

```{r}
# Travel time to work and departure/arrival times
tt <- get_acs_commuting_travel_time(2024, "acs5", "state", state = "MA")
tt$travel_time_residence
tt$departure_time_residence

# Place of work geography
pow <- get_acs_place_of_work(2024, "acs5", "state", state = "MA")
pow$place_of_work_county
pow$workplace_worker_pop

# Transportation mode cross-tabulations (earnings, vehicles available, etc.)
comm_char <- get_acs_commuting_characteristics(2024, "acs5", "state", state = "MA")
comm_char$commuting_earnings_residence
comm_char$commuting_vehicles_residence
```

## Geographical Mobility & Migration

```{r}
# Migration and mobility by current residence
mig_curr <- get_acs_migration_current(2024, "acs5", "state", state = "MA")
mig_curr$migration_age
mig_curr$migration_income

# Migration and mobility by residence 1 year ago
mig_prior <- get_acs_migration_prior(2024, "acs5", "state", state = "MA")
mig_prior$migration_age
mig_prior$migration_income
```

Every component retains `variable`, `label`, `concept`, `table`, `year`, and
`survey`. Derived shares additionally identify their denominator variable.
